---
title: "World Cup"
author: "MsCA Sports Analytics"
date: "`r format(Sys.time(), '%B, %d %Y')`"
output:
  pdf_document:
  github_document:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: inline
always_allow_html: yes
---

```{r Global-options, include=FALSE}
knitr::opts_chunk$set(fig.width=6, fig.height=4, fig.path='Figs/',
                      warning=FALSE, message=FALSE)
```

```{r Preamble, echo=FALSE}
# Enter package in p_load()
# If package is not installed, p_load() will install and load the package
if(!"pacman" %in% rownames(installed.packages())) {
  install.packages("pacman")
  }
pacman::p_load(tidyverse, ggthemes, here, jsonlite)

options(stringsAsFactors = FALSE)

# Set default ggplot theme to tufte
theme_set(ggthemes::theme_tufte())
```

```{r Copy-files, echo=FALSE, eval=FALSE}
# Enter files to load to project directory in from = "~/Downloads/your_file_name_here"
file.copy(from = "~/Downloads/", to = here::here(), 
          overwrite = TRUE, recursive = FALSE, 
          copy.mode = TRUE)
```

# Overview

Models will be built to focus on the following tasks:

* Expected goals: all passes that end in a shot

* Expected possession: all possession for a team

* Pass map

# Data

All the data for the three models exists in `data/events/19714.json`. `19714` represents one game. The events folder has events for all games. There are corresponding `matches` and `lineups` data sets that tie back to the events. We will focus primarily on `events` for now

```{r Data}
events_json <- fromJSON("data/events/19714.json", simplifyVector = FALSE)
```

Parse `events_json` to extract relevant data for **expected goals** model.
```{r Extract-data}
replace_na_empty <- function(x) {
  map_if(x, is.null, ~ NA)
}

event_id <- map_chr(events_json, ~ .x$id)

type_id <- map_int(events_json, ~ .x$type$id)

type_name <- map_chr(events_json, ~ .x$type$name)

timestamp <- map_chr(events_json, ~ .x$timestamp)

possession_team_name <- map_chr(events_json, ~ .x$possession_team$name)

team_name <- map_chr(events_json, ~ .x$team$name)

pass_length <- map(events_json, ~ .x$pass$length) %>% 
  replace_na_empty() %>% unlist()

pass_height <- map(events_json, ~ .x$pass$height$name) %>% 
  replace_na_empty() %>% unlist()

pass_angle <- map(events_json, ~ .x$pass$angle) %>% 
  replace_na_empty() %>% unlist()

duration <- map(events_json, ~ .x$duration) %>% 
  replace_na_empty() %>% unlist()

play_pattern_name <- map_chr(events_json, ~ .x$play_pattern$name)

goalkeeper_type_name <- map(events_json, ~ .x$goalkeeper$type$name) %>% 
  replace_na_empty() %>% unlist()

goalkeeper_outcome_name <- map(events_json, ~ .x$goalkeeper$outcome$name) %>% 
  replace_na_empty() %>% unlist()
```

```{r events-df}
events_df <- data.frame(
    event_id,
    type_id,
    type_name,
    timestamp,
    duration,
    team_name,
    possession_team_name,
    play_pattern_name,
    pass_length,
    pass_height,
    pass_angle,
    goalkeeper_type_name,
    goalkeeper_outcome_name
  ) %>%
  #' Used to identify sequences; max `FALSE` value is the start of a sequence
  mutate(lead_possessor = possession_team_name == lead(possession_team_name)) %>% 
  as_tibble()
```


```{r related-events-df, eval=FALSE, echo=FALSE}
related_events_df <- set_names(map(events_json, ~ .x$related_events), event_id) %>% 
  enframe()

related_events_list <- related_events_df$value %>% map(., function(x) {
  if (is.null(x)) {
    "no event"
  } else {
    x
  }
})

related_events_count <- map(related_events_list, ~ length(.x)) %>% unlist()
related_events <- related_events_list %>% unlist()

related_events_df <- data.frame(main_event_id = rep(event_id, related_events_count), related_event_id = related_events)

related_events_test <- related_events_df %>%
  left_join(events_df, by = c("main_event_id" = "event_id")) %>% 
  left_join(events_df, by = c("related_event_id" = "event_id"), suffix = c("_main_event", "_related_event"))

related_events_test %>% 
  filter(str_detect(type_name_main_event, "Pass"))

related_events_test %>% 
  semi_join(events_df %>% filter(str_detect(type_name, "Shot")), by = c("main_event_id" = "event_id")) %>% 
  count(type_name_main_event, type_name_related_event)
```

Check index of all shots and then look back to see what lead to a shot.
```{r}
(shot_indexes <- which(str_detect(events_df$type_name, "Shot")))
```

Pass sequences are defined as uninterrupted possession leading to a shot.
```{r pass-sequences}
sequence_indexes <- vector("list", length(shot_indexes))

for (i in seq_along(shot_indexes)) {
  start_index <- ifelse(i == 1, 1, shot_indexes[i - 1] + 2)
  sequence_indexes[[i]] <- seq(start_index, shot_indexes[i] + 1, 1)
}

shots_split <- map(sequence_indexes, ~ events_df %>% slice(min(.x):max(.x)))

start_sequence <- map_int(shots_split, function(x) {
  x <- x %>% mutate(type_name_flag = ifelse(lag(type_name) == "Shot", "remove", "keep"))
  x <- x %>% filter(type_name_flag == "keep")
  #' Check if `FALSE` exist and return max index of `FALSE`
  if (length(which(!x$lead_possessor)) >= 1) {
    as.integer(max(which(!x$lead_possessor)) + 2)
  } else {
      min(x$lead_possessor)
    }
  }
) 

pass_sequences <- map2(shots_split, start_sequence, ~ .x %>% slice(.y:nrow(.x))) %>%
  map2(., 1:length(shots_split), ~ mutate(.x, pass_sequence_label = .y)) %>%
  bind_rows() %>%
  mutate(
    pass_sequence_label = factor(pass_sequence_label),
    type_name = ifelse(type_name == "Goal Keeper", goalkeeper_type_name, type_name)
  ) %>%
  group_by(pass_sequence_label) %>%
  # Need to account for shots that were block by someone 
  # other than the goal keeper when identifying outcome of shot
  mutate(goal = ifelse(
    str_detect(
      goalkeeper_outcome_name,
      "Goal Conceded|Penalty Conceded|No Touch|Touched In"
      ), "goal conceded", "saved"
  ))

pass_sequences %>% 
  select(pass_sequence_label, type_name, goalkeeper_outcome_name, goal) %>% 
  mutate_at(vars(goal), ~ ifelse(is.na(.x), "saved", .x)) %>% 
  distinct(pass_sequence_label, goal)
```

```{r Show-data}
pass_sequences %>% 
  group_by(pass_sequence_label) %>% 
  slice(1:6) %>% 
  select(pass_sequence_label, everything())
```

```{r Plot}
pass_sequences %>% 
  group_by(pass_sequence_label) %>% 
  count() %>% 
  ggplot(aes(fct_rev(fct_reorder(pass_sequence_label, n)), n)) +
  geom_col() +
  labs(title = "Event Count by Pass Sequence",
       x = "Pass Sequence",
       y = "Count")

pass_sequences %>% 
  count(pass_sequence_label, type_name) %>% 
  filter(type_name == "Pass") %>% 
  ggplot(aes(fct_rev(fct_reorder(pass_sequence_label, n)), n)) +
  geom_col() +
  labs(title = "Pass Count by Pass Sequence",
       x = "Pass Sequence",
       y = "Count")
```















